Popularizing linear and nonlinear global approaches to hydrodynamic instabilities : A review and a simple implementation for the wake of a cylinder

### 2017-10-10 ARTICLE_STABFEM/ARTICLE_ASME.tex
adding WNL a la Sipp & Lebedev

In the next section we thus present this model and propose a direct resolution method. The relation with the initial method of Mantic-Lugo et al. will be discussed afterwards.

@@

In this section, to simplify the notations, The Navier-Stokes equations will be written symbolically as 
$\partial_t  {\bf u }  = NS(  {\bf u } )$, therefore droping the systematic reference of the incompressibility constraint and associated pressure field.

>>>>

In the next sections we thus successively present the weakly nonlinear model, the self-consistent model and eventually the harmonic-balance model, and propose a direct resolution method. 
%The relation with the initial method of Mantic-Lugo et al. will be discussed afterwards.

In this section, to simplify the notations, The Navier-Stokes equations will be written symbolically as 
$\partial_t  {\bf u }  = NS(  {\bf u } )$, therefore dropping the systematic reference of the incompressibility constraint and associated pressure field.

\paragraph{Weakly nonlinear model}

We give here a brief account of the Weakly nonlinear model of Sipp \& Lebedev, also discussed by
Gallaire et al. (FDR 2016). The derivation is based on a multiple scale decomposition based on the parameter $\epsilon = 1/Re - 1/Re_c$, and the slow time scale $\tau = \epsilon t$. The starting point is an expansion of the flow under the following form :

\begin{eqnarray}
{\bf u} &=& {\bf u}_{bc} + \epsilon \left[ A(\tau) \hat{\bf u}_c e^{i \omega_c t} + c.c. \right] \label{WNL1}\\
&+& \epsilon^2 \left[ {\bf u}_\epsilon + |A|^2  {\bf u}_{2,0} + \left(  A^2 {\bf u}_{2,2} e^{2 i \omega_c t} + c.c. \right) \right] + {\cal O}( \epsilon ^3)
\nonumber
\end{eqnarray}
where ${\bf u}_{bc}$ is the base flow at the threshold $Re_c$, $\hat{\bf u}$ the neutral eigenmode at $Re_c$, $A(\tau)$ a slowly varying amplitude.
The order $\epsilon^2$ contains three terms respectively computed as the solutions of the following linear problems :
\begin{eqnarray}
NSL_{{\bf u}_bc} ({\bf u}_\epsilon) + 2 \nabla \cdot {\bf D}( \hat{\bf u}) &=&0,\\
NSL_{{\bf u}_bc} ({\bf u}_{2,0}) &=& {\cal C}(\hat{\bf u}, ,\overline{\hat{\bf u}}),\\
NSL_{{\bf u}_bc} ({\bf u}_{2,2}) - 2 i \omega_c {\bf u}_{2,2}  &=& \frac{1}{2} {\cal C}(\hat{\bf u},\hat{\bf u}).
 \end{eqnarray}

Finally compatibility conditions are imposed at order 3 leading to an amplitude equation with the following
form :
\be{WNL3}
\frac{\partial A }{\partial \tau} = \Lambda A - (\nu_0+\nu_2)  |A|^2 A
\ee
With 
\begin{eqnarray}
\Lambda &=& \frac{ \left< {\hat{\bf u}}^\dag, 
\left( {\cal C}({\bf u}_\epsilon, \hat{\bf u}) - 2 \nabla \cdot  {\bf D}( \hat{\bf u}) \right) \right>}
{  \left<  \hat{\bf u}^\dag,\hat{\bf u} \right> },\\
\nu_0 &=& -\frac{ \left< {\hat{\bf u}}^\dag,  {\cal C}({\bf u}_{20}, \hat{\bf u} ) \right>}
{  \left<  \hat{\bf u}^\dag,\hat{\bf u} \right> },\\
\nu_2 &=& -\frac{ \left< {\hat{\bf u}}^\dag,  {\cal C}({\bf u}_{22}, \overline{\hat{\bf u}})  \right>}
{  \left<  \hat{\bf u}^\dag,\hat{\bf u} \right> }.
 \end{eqnarray}

If one is interested in the amplitude of the limit cycle, we have 
$|\tilde{A}| = \sqrt{\Lambda_r/(\nu{0,r}+\nu{2,r})}$ and eventually, reintroducing the scaling :
\be 
|A| =  \frac{\Lambda_r}{\nu{0,r}+\nu{2,r} } \sqrt{\frac{1}{Re_c}-\frac{1}{Re}}
\ee

The approach gives good prediction in the imediate vicinity of $Re_c$, but deviation appears very 
rapidly and disagreement is already large at $Re=48$ (although, as discussed by Galaire et al. 
and also observed by Tchoufag et al. (PRL), a convenient definition of the $\epsilon$ improves the results).

@@
